---
title: "Take-home Exercise 1"
description: |
  In this take home exercise, i will explore ......
author:
  - name: Aloysius Teng
    url: https://www.linkedin.com/in/aloysius-teng-32477716a
    affiliation: School of Computing and Information System
    affiliation_url: https://scis.smu.edu.sg/
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
In this take-home exercise, appropriate static statistical graphics methods are reveal the demographic of the city of Engagement, Ohio USA. 
The data was be processed by using appropriate tidyverse family of packages and the statistical graphics was prepared using ggplot2 and its extensions.

# Getting Started

Before we get started, it is important for us to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment.

The chunk code below will do the trick.

```{r}
packages = c('tidyverse', 'ggdist', 'gghalves', 'ggridges')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# Importing Data

The code chunk below imports *Participants.csv* form the data folder into R by using the [`read_csv()`](https://readr.tidyverse.org/reference/read_delim.html) function of [**readr**](https://readr.tidyverse.org/index.html) and saves it as a tibble data frame called *participants*.

```{r}
participants <- read_csv("data/Participants.csv")
```

# A Simple Bar Chart

The code chunk below plots a bar chart by using geom_bar() of ggplot2.

```{r}
ggplot(data=participants,
       aes(x = interestGroup)) +
  geom_bar()
```


